### 2018-04-21 Manuscript/chapter4.tex
Incorporate discussion section suggestions from Dr. Toth

Somewhat surprisingly, the transcriptomic response to virus infection in our experiment was fairly limited. We found only 43 transcripts to be differentially expressed, some with known immune functions (Table \ref{tbl:virusGenes}) such as argonaute-2 and a gene with similarity to MD-2 lipid recognition protein, as well as additional genes related to transcriptional regulation, and muscle contraction.  The small number of DEGs in this study may be partly explained by the large amount of noise in the data (Figures \ref{fig:pcpRutterVirus}, \ref{fig:litreClusterRutter}, \ref{fig:RutterSM1}, \ref{fig:RutterSM2}, \ref{fig:RutterSM3}, \ref{fig:RutterSM4}).

Given the noisy nature of our data, and our desire to hone in on genes with real expression differences, we compared our data to Galbraith et al. (2017), a study that also examined bees’ response to viral infection.  In contrast to our study, Galbraith et al. identified a large number of virus responsive trancripts, and generally had less noise in their data (Figures X, X, X).  To identify the most reliable virus-responsive genes from our study, we looked for overlap in the DEGs associated with virus infection on both experiments.  We found a large, statistically significant overlap, with 26/38 (68%) of virus-responsive DEGs from our study also showing response to virus infection in Galbraith et al. (2017, Figure 1.5).  This result gives us confidence that, although noisy, we were able to uncover consistent, replicable gene expression responses to virus infection.

@@

One of the goals of this study was to use our RNA-seq data to assess whether transcriptomic responses to diet quality and virus infection provide insight into whether high quality diet can buffer bees from pathogen stress via mechanisms of “resistance” or “resilience”.  We attempted to address this through specific gene expression contrasts (Table 1.10, Figure 1.18), accompanied by GO analysis of the associated gene lists.  We an approximately equal number of resistance (n=122) and resilience (n=125) related candidate genes, suggesting both processes may be playing significant roles in dietary buffering from pathogen infection.  Resilience candidate genes had functions related to carbohydrate metabolism, chitin metabolism, immune response, and regulation of transcription.  Resistance candidate genes had functions related to several forms of metabolism (chitin, carbohydrate), regulation of transcription, and cell adhesion.

@@

1) Diet effect seemed larger by number of DEGs than virus effect \\

2) No interaction terms in DEGs \\

3) Elaborate more on the GOs that came out of different contrasts \\

4) Possible preliminary evidence for resilience effects, whereby immune-related genes are turned on by good diet in preparation to defend against viral attacks \\

5) Galbraith data was much cleaner as expected due to 75 percent genetic similarity. However, we had significant overlap in the virus-related DEGs. Possible avenue for future studies to explore as the intersection DEGs between these two studies may represent those that are reliable (well-controlled study) and relevant (study conducted in more realistic environmental settings).

>>>>

Somewhat surprisingly, the transcriptomic response to virus infection in our experiment was fairly limited. We found only 43 transcripts to be differentially expressed, some with known immune functions (Table \ref{tbl:virusGenes}) such as argonaute-2 and a gene with similarity to MD-2 lipid recognition protein, as well as additional genes related to transcriptional regulation, and muscle contraction. The small number of DEGs in this study may be partly explained by the large amount of noise in the data (Figures \ref{fig:mdsPlots}B, \ref{fig:pcpRutterVirus}, \ref{fig:litreClusterRutter}, \ref{fig:RutterSM1}, \ref{fig:RutterSM2}, \ref{fig:RutterSM3}, \ref{fig:RutterSM4}).

Given the noisy nature of our data, and our desire to hone in on genes with real expression differences, we compared our data to the Galbraith study (\citealt{galbraith}), which also examined bees response to viral infection. In contrast to our study, Galbraith et al. identified a large number of virus responsive trancripts, and generally had less noise in their data (Figures \ref{fig:mdsPlots}A, \ref{fig:pcpGalbraith}, \ref{fig:litreCluster2}, \ref{fig:litreCluster1}, \ref{fig:GalbraithClust1SM}, \ref{fig:GalbraithClust2SM}, \ref{fig:GalbraithClust3SM}, \ref{fig:GalbraithClust4SM}).  To identify the most reliable virus-responsive genes from our study, we looked for overlap in the DEGs associated with virus infection on both experiments. We found a large, statistically significant (p-value < 2.2e-16) overlap, with 26/38 (68\%) of virus-responsive DEGs from our study also showing response to virus infection in Galbraith et al. (Figure \ref{fig:GRVenn}). This result gives us confidence that, although noisy, we were able to uncover consistent, replicable gene expression responses to virus infection with our data.

@@

One of the goals of this study was to use our RNA-seq data to assess whether transcriptomic responses to diet quality and virus infection provide insight into whether high quality diet can buffer bees from pathogen stress via mechanisms of ``resistance'' or ``resilience''. We attempted to address this question through specific gene expression contrasts (Table \ref{tbl:contrasts}, Figure \ref{tbl:NCNRNCPathways}), accompanied by GO analysis of the associated gene lists. We found an approximately equal number of resistance (n = 122) and resilience (n = 125) related candidate genes, suggesting both processes may be playing significant roles in dietary buffering from pathogen infection. Resilience candidate genes had functions related to carbohydrate metabolism, chitin metabolism, immune response, and regulation of transcription. Resistance candidate genes had functions related to several forms of metabolism (chitin and carbohydrate), regulation of transcription, and cell adhesion.

@@
